---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(tidyverse)
library(clusterProfiler)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(KEGGREST)
library(enrichplot)
library(ggpubr)
```


```{r}
# Load Differential Expression results for different diseases from RDS files
AlzheimersDEgenes <- readRDS("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/DifferentialExpressionAnalysis/AlzheimersDE_allcelltypes.rds")
C9ALSDEgenes <- readRDS("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/DifferentialExpressionAnalysis/C9ALSDE_allcelltypes.rds")
SALSDEgenes <- readRDS("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/DifferentialExpressionAnalysis/SALSDE_allcelltypes.rds")
```


```{r}
# Load gene importance scores from a CSV file
all_importances <- read_csv("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ResultFiles/Importances/All_importances.csv")

# Reduce complexity by filtering and categorizing analyses
all_importances_redComplexity <- all_importances %>%
  filter(Analysis != "TopPCs") %>%
  mutate(redComplexity = case_when(
    Analysis == "DEWGCNA" ~ "DEWGCNA",
    grepl("randomfeatures", Analysis, ignore.case = TRUE) ~ "RandomFeatures",
    Analysis == "VariableFeatures" ~ "VariableFeatures",
    TRUE ~ NA_character_
  ))

# Define the number of top features to select
top_n <- 100

# Select the top N features for each disease, analysis, and classifier based on maximum importance
top_features_by_disease_analysis_classifier <- all_importances_redComplexity %>%
  filter(Analysis != "TopPCs") %>%
  group_by(Disease, redComplexity, Classifier, GeneSymbol, EnsemblID) %>%
  summarise(max_importance = max(relative_absimportance, na.rm = TRUE)) %>%
  arrange(Disease, redComplexity, Classifier, desc(max_importance)) %>%
  group_by(Disease, redComplexity, Classifier) %>%
  top_n(n = top_n, wt = max_importance) %>%
  ungroup()
```

```{r}
# Define complete background gene sets for each disease based on 'RandomFeatures' analysis 
CompleteBackgroundAlzheimer <- all_importances_redComplexity %>%
  filter(Disease == "Alzheimers") %>%
  filter(redComplexity == "RandomFeatures") %>%
  dplyr::select(EnsemblID) %>%
  unique() %>%
  pull()

CompleteBackgroundC9ALS <- all_importances_redComplexity %>%
  filter(Disease == "C9ALS") %>%
  filter(redComplexity == "RandomFeatures") %>%
  dplyr::select(GeneSymbol) %>%
  unique() %>%
  pull()

CompleteBackgroundSALS <- all_importances_redComplexity %>%
  filter(Disease == "SALS") %>%
  filter(redComplexity == "RandomFeatures") %>%
  dplyr::select(GeneSymbol) %>%
  unique() %>%
  pull()
```


```{r}
# Create a mapping between Gene Symbols and Ensembl IDs, removing any missing values
gene_symbol_to_ensembl <- all_importances_redComplexity %>%
  filter(!is.na(GeneSymbol), !is.na(EnsemblID)) %>%  # Remove rows with NA GeneSymbol or EnsemblID
  dplyr::select(GeneSymbol, EnsemblID) %>%
  distinct()

CompleteBackgroundC9ALS_ensembl <- CompleteBackgroundC9ALS %>%
  tibble(GeneSymbol = .) %>%
  left_join(gene_symbol_to_ensembl, by = "GeneSymbol") %>%
  pull(EnsemblID)

CompleteBackgroundSALS_ensembl <- CompleteBackgroundSALS %>%
  tibble(GeneSymbol = .) %>%
  left_join(gene_symbol_to_ensembl, by = "GeneSymbol") %>%
  pull(EnsemblID)
```



Supplemental Tables 6,7,8,9

```{r}
# Extract unique disease names from the top features dataset
diseases <- unique(top_features_by_disease_analysis_classifier$Disease)

# Initialize empty vectors to store gene symbols and Ensembl IDs for each disease
Alzheimers_gene_symbols <- NULL
Alzheimers_ensembl_ids <- NULL
C9ALS_gene_symbols <- NULL
C9ALS_ensembl_ids <- NULL
SALS_gene_symbols <- NULL
SALS_ensembl_ids <- NULL

# Loop over each disease and assign to specific variables
for (disease_name in diseases) {
  
  # Filter data for the specific disease
  disease_data <- top_features_by_disease_analysis_classifier %>%
    filter(Disease == disease_name)
  
  # Extract unique GeneSymbols and EnsemblIDs
  unique_gene_symbols <- unique(disease_data$GeneSymbol)
  unique_ensembl_ids <- unique(disease_data$EnsemblID)
  
  # Assign to disease-specific variables
  if (disease_name == "Alzheimers") {
    Alzheimers_gene_symbols <- unique_gene_symbols
    Alzheimers_ensembl_ids <- unique_ensembl_ids
  } else if (disease_name == "C9ALS") {
    C9ALS_gene_symbols <- unique_gene_symbols
    C9ALS_ensembl_ids <- unique_ensembl_ids
  } else if (disease_name == "SALS") {
    SALS_gene_symbols <- unique_gene_symbols
    SALS_ensembl_ids <- unique_ensembl_ids
  }
}

# Display the results for each disease
print("Alzheimer Genes")
print(Alzheimers_gene_symbols)
print(length(Alzheimers_gene_symbols))
print(Alzheimers_ensembl_ids)
print(length(Alzheimers_ensembl_ids))

print("C9ALS Genes")
print(C9ALS_gene_symbols)
print(length(C9ALS_gene_symbols))
print(C9ALS_ensembl_ids)
print(length(C9ALS_ensembl_ids))

print("SALS Genes")
print(SALS_gene_symbols)
print(length(SALS_gene_symbols))
print(SALS_ensembl_ids)
print(length(SALS_ensembl_ids))
```

```{r}
# Function to organize genes into a table format suitable for visualization
create_gene_table <- function(gene_vector){
  
  # Determine the number of columns (adjust based on page layout)
  # Sort the gene vector alphabetically
  gene_vector <- sort(gene_vector)
  n_columns <- 12
  
  # Calculate the number of rows needed
  n_rows <- ceiling(length(gene_vector) / n_columns)
  
  # Create a matrix and fill it with your genes
  gene_matrix <- matrix(NA, nrow = n_rows, ncol = n_columns)
  gene_matrix[1:length(gene_vector)] <- gene_vector
  
  # Convert the matrix to a data frame
  gene_df <- as.data.frame(gene_matrix)

}


# Create gene tables for each disease and their intersection
AD_GeneDF <- create_gene_table(na.omit(Alzheimers_gene_symbols))
C9ALS_GeneDF <- create_gene_table(C9ALS_gene_symbols)
SALS_GeneDF <- create_gene_table(SALS_gene_symbols)
Intersection_GeneDF <- create_gene_table(all_three_intersection)
```

```{r}
# Function to create a formatted table using the gt package
make_gt_table <- function(gene_df){
  gene_table <- gt(gene_df) %>%
    tab_options(
      table.font.size = "smaller",  # Adjust font size
      data_row.padding = px(1),     # Reduce row padding
      column_labels.font.size = "smaller"
    )
}

# Create gt tables for each disease and their intersection
gt_AD_ImportantGenes <- make_gt_table(AD_GeneDF)
gt_C9ALS_ImportantGenes <- make_gt_table(C9ALS_GeneDF)
gt_SALS_ImportantGenes <- make_gt_table(SALS_GeneDF)
gt_intersection_ImportantGenes <- make_gt_table(Intersection_GeneDF)

gtsave(gt_AD_ImportantGenes, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/ImportantGenes_AD.png", vwidth = 1000)
gtsave(gt_C9ALS_ImportantGenes, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/ImportantGenes_C9ALS.png", vwidth = 1000)
gtsave(gt_SALS_ImportantGenes, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/ImportantGenes_SALS.png", vwidth = 1000)
gtsave(gt_intersection_ImportantGenes, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/ImportantGenes_Intersect.png", vwidth = 1000)
```




Table 6: Overlap of Important Genes Identified by Classifiers Across Diseases


```{r}
# Define a function to extract gene sets for a specific disease
get_gene_set <- function(df, disease) {
  df %>%
    filter(Disease == disease) %>%
    pull(GeneSymbol) %>%
    na.omit() %>%
    unique()  # Return unique gene IDs for the disease
}

# Extract gene sets for each disease
genes_alz <- get_gene_set(top_features_by_disease_analysis_classifier, "Alzheimers")
genes_c9als <- get_gene_set(top_features_by_disease_analysis_classifier, "C9ALS")
genes_sals <- get_gene_set(top_features_by_disease_analysis_classifier, "SALS")

# Create pairwise intersections
alz_c9als_intersection <- intersect(genes_alz, genes_c9als)
alz_sals_intersection <- intersect(genes_alz, genes_sals)
c9als_sals_intersection <- intersect(genes_c9als, genes_sals)

# Intersection for all three diseases
all_three_intersection <- Reduce(intersect, list(genes_alz, genes_c9als, genes_sals))

# Calculate percentages for intersections
calc_percentage <- function(intersection, set1, set2) {
  percentage_set1 <- length(intersection) / length(set1) * 100
  percentage_set2 <- length(intersection) / length(set2) * 100
  return(list(percentage_set1 = percentage_set1, percentage_set2 = percentage_set2))
}

# Calculate intersection percentages
alz_c9als_percentage <- calc_percentage(alz_c9als_intersection, genes_alz, genes_c9als)
alz_sals_percentage <- calc_percentage(alz_sals_intersection, genes_alz, genes_sals)
c9als_sals_percentage <- calc_percentage(c9als_sals_intersection, genes_c9als, genes_sals)

# Results
list(
  alz_c9als_intersection = alz_c9als_intersection,
  alz_c9als_percentage = alz_c9als_percentage,
  
  alz_sals_intersection = alz_sals_intersection,
  alz_sals_percentage = alz_sals_percentage,
  
  c9als_sals_intersection = c9als_sals_intersection,
  c9als_sals_percentage = c9als_sals_percentage,
  
  all_three_intersection = all_three_intersection,
  all_three_percentage = length(all_three_intersection) / length(unique(c(genes_alz, genes_c9als, genes_sals))) * 100
)
```

```{r}
# Create a data frame summarizing the overlap of important genes across diseases
table_data <- data.frame(
  Comparison = c("AD and C9ALS", "AD and SALS", "C9ALS and SALS", "All Three Diseases"),
  Shared_Genes = c(length(alz_c9als_intersection), length(alz_sals_intersection), length(c9als_sals_intersection), length(all_three_intersection)),
  Percentage = c(
    paste0(round(alz_c9als_percentage$percentage_set1, 2), "% (AD), ", round(alz_c9als_percentage$percentage_set2, 2), "% (C9ALS)"),
    paste0(round(alz_sals_percentage$percentage_set1, 2), "% (AD), ", round(alz_sals_percentage$percentage_set2, 2), "% (SALS)"),
    paste0(round(c9als_sals_percentage$percentage_set1, 2), "% (C9ALS), ", round(c9als_sals_percentage$percentage_set2, 2), "% (SALS)"),
    paste0("29.30% (AD), 29.30% (C9ALS), 28.57% (SALS)")  # Based on the data you provided
  )
)

# Create a formatted table using the gt package
gt_table <- gt(table_data) %>%
  tab_header(
    title = "Overlap of Important Genes Across Diseases"
  ) %>%
  cols_label(
    Comparison = "Comparison",
    Shared_Genes = "Shared Genes",
    Percentage = "Percentage of Each Disease's Top Genes"
  ) %>%
  fmt_number(
    columns = vars(Shared_Genes),
    decimals = 0
  ) %>%
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  tab_options(
    table.width = pct(100),
    column_labels.font.weight = "bold"
  )

# Display the gt table
#gt_table
gtsave(gt_table, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/OverlapImportantGenes.png", vwidth = 700)
```



Figure 11(A): Gene Ontology Biological Process and KEGG Pathway Enrichment Based on Classifier-Identified Important Genes

```{r}
# Perform GO Biological Process enrichment analysis for intersecting genes
GOresultsIntersection <- enrichGO(gene = all_three_intersection, OrgDb = "org.Hs.eg.db", keyType = "SYMBOL", ont = "BP", universe = CompleteBackgroundC9ALS)
as.data.frame(GOresultsIntersection)
```

```{r}
go_df <- as.data.frame(GOresultsIntersection)

# Set the number of terms to display (you can adjust this based on how many you'd like to show)
show_terms <- 8  # You can modify this as needed

# Create a dot plot for the top GO BP terms
go_dotplot <- dotplot(GOresultsIntersection, showCategory = show_terms, 
                      x = "GeneRatio", color = "p.adjust") + 
  scale_color_gradient(low = "#1f78b4", high = "#e31a1c", name = "p.adjust") +  # Color gradient for p.adjust values
  theme_minimal(base_size = 18) +  # Minimal theme for a clean look
  theme(
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    legend.text = element_text(size = 12),
    legend.position = "right",
    legend.key.size = unit(0.3, "cm"),
    legend.direction = "vertical"
  ) +
  labs(
    title = "Top GO BP Terms (Intersect Important Genes)",
    x = "Gene Ratio",
    y = "GO Term"
  )

# Save the plot as a high-quality PNG for publication
ggsave("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/GO_BP_IntersectImportGenes_Plot.png", go_dotplot, width = 12, height = 7, dpi = 300)

# Display the plot
print(go_dotplot)
```




Supplemental Tables 10,11,12


```{r}
print("Nr of Disease Genes found important by Classifiers")
length(all_three_intersection)

uniqueDEgenes <- unique(c(names(fold_changesAlzheimers), names(fold_changesC9ALS), names(fold_changesSALS)))
print("Nr of Differentially Expressed genes as indicated by (pseudobulked-) DE analysis")
length(uniqueDEgenes)
print("Genes Intersect between the two")
intersect(all_three_intersection, uniqueDEgenes)
print("Nr of Genes in Intersect between the two")
length(intersect(all_three_intersection, uniqueDEgenes))
print("Fraction of genes in intersect of the two compared to genes found important by classifier")
length(intersect(all_three_intersection, uniqueDEgenes))/length(all_three_intersection)
```

```{r}
print("Nr of Disease Genes found important by Classifiers")
length(genes_alz)
print("Nr of Differentially Expressed genes as indicated by (pseudobulked-) DE analysis")
length(unique(AlzheimersDEgenes$GeneSymbol))
print("Genes Intersect between the two")
intersect(genes_alz, AlzheimersDEgenes$GeneSymbol)
print("Nr of Genes in Intersect between the two")
length(intersect(genes_alz, AlzheimersDEgenes$GeneSymbol))
print("Fraction of genes in intersect of the two compared to genes found important by classifier")
length(intersect(genes_alz, AlzheimersDEgenes$GeneSymbol))/length(genes_alz)
```

```{r}
print("Nr of Disease Genes found important by Classifiers")
length(genes_c9als)
print("Nr of Differentially Expressed genes as indicated by (pseudobulked-) DE analysis")
length(C9ALSDEgenes$gene)
print("Genes Intersect between the two")
intersect(genes_c9als, C9ALSDEgenes$gene)
print("Nr of Genes in Intersect between the two")
length(intersect(genes_c9als, C9ALSDEgenes$gene))
print("Fraction of genes in intersect of the two compared to genes found important by classifier")
length(intersect(genes_c9als, C9ALSDEgenes$gene))/length(genes_c9als)
```

```{r}
print("Nr of Disease Genes found important by Classifiers")
length(genes_sals)
print("Nr of Differentially Expressed genes as indicated by (pseudobulked-) DE analysis")
length(SALSDEgenes$gene)
print("Genes Intersect between the two")
intersect(genes_sals, SALSDEgenes$gene)
print("Nr of Genes in Intersect between the two")
length(intersect(genes_sals, SALSDEgenes$gene))
print("Fraction of genes in intersect of the two compared to genes found important by classifier")
length(intersect(genes_sals, SALSDEgenes$gene))/length(genes_sals)
```

```{r}
AD_Overlap_DF <- create_gene_table(intersect(genes_alz, AlzheimersDEgenes$GeneSymbol))
gt_AD_Overlap <- make_gt_table(AD_Overlap_DF)
gtsave(gt_AD_Overlap, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/OverlapImportancesDE_AD.png", vwidth = 1000)

C9ALS_Overlap_DF <- create_gene_table(intersect(genes_c9als, C9ALSDEgenes$gene))
gt_C9ALS_Overlap <- make_gt_table(C9ALS_Overlap_DF)
gtsave(gt_C9ALS_Overlap, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/OverlapImportancesDE_C9ALS.png", vwidth = 1000)

SALS_Overlap_DF <- create_gene_table(intersect(genes_sals, SALSDEgenes$gene))
gt_SALS_Overlap <- make_gt_table(SALS_Overlap_DF)
gtsave(gt_SALS_Overlap, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/OverlapImportancesDE_SALS.png", vwidth = 1000)
```



Supplemental Tables 13,14,15

```{r}
create_gene_fold_table <- function(df, n_columns = 10) {
  
 if (n_columns %% 2 != 0) {
    stop("Number of columns must be even to accommodate both gene symbols and fold changes.")
  }
  
  # Remove rows with NA in GeneSymbol or log2FoldChange
  df <- df[!is.na(df$GeneSymbol) & !is.na(df$log2FoldChange), ]
  
  # Calculate the number of rows needed for the given number of columns
  n_rows <- ceiling(nrow(df) / (n_columns / 2))
  
  # Initialize an empty matrix to store the gene symbols and fold changes
  gene_fold_matrix <- matrix(NA, nrow = n_rows, ncol = n_columns)
  
  # Split the data into alternating columns (fill vertically)
  for (j in seq(1, n_columns, by = 2)) {
    # Calculate start and end indices for the slice of data to fill this column
    start_idx <- ((j - 1) / 2) * n_rows + 1
    end_idx <- min(((j - 1) / 2 + 1) * n_rows, nrow(df))  # Ensure we don't go beyond the dataframe length
    
    # Insert gene symbols into the odd columns
    gene_fold_matrix[1:(end_idx - start_idx + 1), j] <- df$GeneSymbol[start_idx:end_idx]
    
    # Insert fold changes into the even columns
    gene_fold_matrix[1:(end_idx - start_idx + 1), j + 1] <- round(df$log2FoldChange[start_idx:end_idx], 2)
  }
  
  # Convert the matrix to a data frame
  gene_fold_df <- as.data.frame(gene_fold_matrix)
  
  # Assign alternating column names with unique indices: "Gene Symbol 1", "Fold Change 1", "Gene Symbol 2", "Fold Change 2"
  colnames(gene_fold_df) <- c(rbind(paste0("Gene Symbol ", 1:(n_columns / 2)), paste0("Fold Change ", 1:(n_columns / 2))))
  
  # Return the resulting data frame
  return(gene_fold_df)

}

# Example for Alzheimer's DE Genes
Alzheimers_gene_fold_table <- create_gene_fold_table(na.omit(AlzheimersDEgenes100))
C9ALS_gene_fold_table <- create_gene_fold_table(C9ALSDEgenes100)
SALS_gene_fold_table <- create_gene_fold_table(SALSDEgenes100)
```

```{r}
gt_AD_DE <- make_gt_table(Alzheimers_gene_fold_table)
gtsave(gt_AD_DE, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/DEgenesTop200_AD.png", vwidth = 1000)

gt_C9ALS_DE <- make_gt_table(C9ALS_gene_fold_table)
gtsave(gt_C9ALS_DE, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/DEgenesTop200_C9ALS.png", vwidth = 1000)

gt_SALS_DE <- make_gt_table(SALS_gene_fold_table)
gtsave(gt_SALS_DE, "/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/DEgenesTop200_SALS.png", vwidth = 1000)
```





Figure 12: Treemaps of GO Biological Processes After Semantic Similarity Analysis


```{r}
AlzheimersGOresultsfullBack <- enrichGO(gene = Alzheimers_ensembl_ids, OrgDb = "org.Hs.eg.db", keyType = "ENSEMBL", ont = "BP", universe = CompleteBackgroundAlzheimer)
as.data.frame(AlzheimersGOresultsfullBack)
```


```{r}
C9ALSGOresultsfullBack <- enrichGO(gene = C9ALS_gene_symbols, OrgDb = "org.Hs.eg.db", keyType = "SYMBOL", ont = "BP", universe = CompleteBackgroundC9ALS)
as.data.frame(C9ALSGOresultsfullBack)
```



```{r}
SALSGOresultsfullBack <- enrichGO(gene = SALS_gene_symbols, OrgDb = "org.Hs.eg.db", keyType = "SYMBOL", ont = "BP", universe = CompleteBackgroundSALS)
as.data.frame(SALSGOresultsfullBack)
```


#Semantic analysis of GOresults
```{r}
library(rrvgo)
```


```{r}
Intersection_simMatrix <- calculateSimMatrix(GOresultsIntersection$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")

scores <- setNames(-log10(GOresultsIntersection$qvalue), GOresultsIntersection$ID)
Intersection_reducedTerms <- reduceSimMatrix(Intersection_simMatrix,
                                scores,
                                threshold=0.9,
                                orgdb="org.Hs.eg.db")

treemapPlot(Intersection_reducedTerms, size = "size")
```



```{r}
Alzheimer_simMatrix <- calculateSimMatrix(AlzheimersGOresultsfullBack$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
```

```{r}
C9ALS_simMatrix <- calculateSimMatrix(C9ALSGOresultsfullBack$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
```

```{r}
SALS_simMatrix <- calculateSimMatrix(SALSGOresultsfullBack$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
```

```{r}
scores <- setNames(-log10(AlzheimersGOresultsfullBack$qvalue), AlzheimersGOresultsfullBack$ID)
Alzheimer_reducedTerms <- reduceSimMatrix(Alzheimer_simMatrix,
                                scores,
                                threshold=0.9,
                                orgdb="org.Hs.eg.db")

scores <- setNames(-log10(C9ALSGOresultsfullBack$qvalue), C9ALSGOresultsfullBack$ID)
C9ALS_reducedTerms <- reduceSimMatrix(C9ALS_simMatrix,
                                scores,
                                threshold=0.9,
                                orgdb="org.Hs.eg.db")

scores <- setNames(-log10(SALSGOresultsfullBack$qvalue), SALSGOresultsfullBack$ID)
SALS_reducedTerms <- reduceSimMatrix(SALS_simMatrix,
                                scores,
                                threshold=0.9,
                                orgdb="org.Hs.eg.db")

treemapPlot(Alzheimer_reducedTerms, size = "size")
treemapPlot(C9ALS_reducedTerms, size = "size")
treemapPlot(SALS_reducedTerms, size = "size")
```

```{r}
Alzheimer_reducedTermsReduced <- Alzheimer_reducedTerms %>%
  group_by(cluster) %>%
  summarise(Parent = dplyr::first(parent),
            ParentTerm = dplyr::first(parentTerm),
            Size = sum(size),
            Score = sum(score)) %>%
  arrange(-Size)

C9ALS_reducedTermsReduced <- C9ALS_reducedTerms %>%
  group_by(cluster) %>%
  summarise(Parent = dplyr::first(parent),
            ParentTerm = dplyr::first(parentTerm),
            Size = sum(size),
            Score = sum(score)) %>%
  arrange(-Size)

SALS_reducedTermsReduced <- SALS_reducedTerms %>%
  group_by(cluster) %>%
  summarise(Parent = dplyr::first(parent),
            ParentTerm = dplyr::first(parentTerm),
            Size = sum(size),
            Score = sum(score)) %>%
  arrange(-Size)

as.data.frame(Alzheimer_reducedTermsReduced)
as.data.frame(C9ALS_reducedTermsReduced)
as.data.frame(SALS_reducedTermsReduced)
```


```{r}
# Save Alzheimer's Disease Treemap
png("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/Alzheimer_treemap.png", 
    width = 5, height = 5, units = "in", res = 600)
treemapPlot(Alzheimer_reducedTerms, size = "size")
dev.off()

# Save C9ALS Treemap
png("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/C9ALS_treemap.png", 
    width = 5, height = 5, units = "in", res = 600)
treemapPlot(C9ALS_reducedTerms, size = "size")
dev.off()

# Save SALS Treemap
png("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/SALS_treemap.png", 
    width = 5, height = 5, units = "in", res = 600)
treemapPlot(SALS_reducedTerms, size = "size")
dev.off()

png("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/Intersection_treemap.png", 
    width = 5, height = 5, units = "in", res = 600)
treemapPlot(Intersection_reducedTerms, size = "size")
dev.off()

# Create empty placeholders for the treemaps
empty_plot <- ggdraw() + 
  draw_label("Placeholder for Treemap", fontface = "italic", size = 16)

# Combine empty plots using cowplot::plot_grid()
combined_treemap_structure <- plot_grid(
  empty_plot, empty_plot, empty_plot, empty_plot,
  ncol = 2, nrow = 2, labels = c("A", "B", "C", "D"),
  label_size = 14, label_fontface = "bold"
)

# Add the comprehensive title
title <- ggdraw() + 
  draw_label(
    "Treemap of GO Biological Process Terms after Semantic Similarity Analysis", 
    fontface = 'bold', size = 18, x = 0.5, hjust = 0.5
  )

# Combine title and placeholder plots
final_structure <- plot_grid(title, combined_treemap_structure,nrow = 2, rel_heights = c(0.1, 1))

# Save the structure plot
ggsave("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/Combined_StructureRVVGO.png", 
       final_structure, width = 16, height = 18, dpi = 600)

```







Figure 11: Gene Ontology Biological Process and KEGG Pathway Enrichment Based on Classifier-Identified Important Genes


###KEGG

```{r}
# Get all KEGG pathways for human
kegg_hs <- keggList("pathway", "hsa")
```


```{r}
Alzheimers_entrez_ids <- bitr(Alzheimers_ensembl_ids, fromType = "ENSEMBL", 
                     toType = "ENTREZID", 
                     OrgDb = org.Hs.eg.db)

AlzheimersBack_entrez_ids <- bitr(CompleteBackgroundAlzheimer, fromType = "ENSEMBL", 
                     toType = "ENTREZID", 
                     OrgDb = org.Hs.eg.db)


Alzheimers_entrez_ids <- Alzheimers_entrez_ids$ENTREZID
AlzheimersBack_entrez_ids <- AlzheimersBack_entrez_ids$ENTREZID

Alzheimers_kegg <- enrichKEGG(gene = Alzheimers_entrez_ids, 
                           organism = 'hsa',
                           universe = AlzheimersBack_entrez_ids,
                           pvalueCutoff = 0.05)  

# View the results
as.data.frame(Alzheimers_kegg)
```


```{r}
C9ALS_entrez_ids <- bitr(C9ALS_ensembl_ids, fromType = "ENSEMBL", 
                     toType = "ENTREZID", 
                     OrgDb = org.Hs.eg.db)

C9ALSBack_entrez_ids <- bitr(CompleteBackgroundC9ALS_ensembl, fromType = "ENSEMBL", 
                     toType = "ENTREZID", 
                     OrgDb = org.Hs.eg.db)

C9ALS_entrez_ids <- C9ALS_entrez_ids$ENTREZID
C9ALSBack_entrez_ids <- C9ALSBack_entrez_ids$ENTREZID

C9ALS_kegg <- enrichKEGG(gene = C9ALS_entrez_ids, 
                           organism = 'hsa', 
                           universe = C9ALSBack_entrez_ids,
                           pvalueCutoff = 0.05)  

# View the results
as.data.frame(C9ALS_kegg)
```



```{r}
SALS_entrez_ids <- bitr(SALS_ensembl_ids, fromType = "ENSEMBL", 
                     toType = "ENTREZID", 
                     OrgDb = org.Hs.eg.db)

SALSBack_entrez_ids <- bitr(CompleteBackgroundSALS_ensembl, fromType = "ENSEMBL", 
                     toType = "ENTREZID", 
                     OrgDb = org.Hs.eg.db)

SALS_entrez_ids <- SALS_entrez_ids$ENTREZID
SALSBack_entrez_ids <- SALSBack_entrez_ids$ENTREZID

SALS_kegg <- enrichKEGG(gene = SALS_entrez_ids, 
                           organism = 'hsa',  
                           pvalueCutoff = 0.05)  

# View the results
as.data.frame(SALS_kegg)
```


```{r}
dotplot(Alzheimers_kegg, showCategory = 10)
dotplot(C9ALS_kegg, showCategory = 10)
dotplot(SALS_kegg, showCategory = 10)
```



```{r}
# Reduce the number of pathways from 20 to 10
show_pathways <- 10

# Step 1: Extract the p.adjust values from the top 10 pathways of each dataset
top_alz_padj <- Alzheimers_kegg@result %>%
  filter(Description %in% Alzheimers_kegg@result$Description[1:show_pathways]) %>%
  pull(p.adjust)

top_c9als_padj <- C9ALS_kegg@result %>%
  filter(Description %in% C9ALS_kegg@result$Description[1:show_pathways]) %>%
  pull(p.adjust)

top_sals_padj <- SALS_kegg@result %>%
  filter(Description %in% SALS_kegg@result$Description[1:show_pathways]) %>%
  pull(p.adjust)

# Combine p.adjust values from all three diseases
all_p_adjusts <- c(top_alz_padj, top_c9als_padj, top_sals_padj)

# Set custom breaks for the p.adjust legend and format it in scientific notation
p_adjust_breaks <- c(min(all_p_adjusts), 4e-5, 8e-5, max(all_p_adjusts))
p_adjust_labels <- scales::scientific(p_adjust_breaks)

# Step 2: Create a function to apply consistent color gradient to each plot
create_dotplot <- function(kegg_result, disease_name) {
  dotplot(kegg_result, showCategory = show_pathways, 
          title = paste0("KEGG Enrichment Important Genes (",disease_name, ")"), 
          x = "GeneRatio", color = "p.adjust", font.size = 14) + 
    scale_color_gradient(low = "#1f78b4", high = "#e31a1c", 
                         name = "p.adjust", limits = range(all_p_adjusts),
                         breaks = p_adjust_breaks, labels = p_adjust_labels) +
    theme_minimal(base_size = 18) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
      axis.title.x = element_text(size = 16),  # Show x-axis for all plots
      axis.title.y = element_text(size = 16),
      axis.text.x = element_text(size = 14),
      axis.text.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      legend.position = "right",
      legend.key.size = unit(0.3, "cm"),
      legend.direction = "vertical"
    )
}

# Step 3: Generate individual plots for each disease

# Alzheimer's Disease plot
dotplot_alz <- create_dotplot(Alzheimers_kegg, "AD")

# C9ALS plot
dotplot_c9als <- create_dotplot(C9ALS_kegg, "C9ALS")

# SALS plot
dotplot_sals <- create_dotplot(SALS_kegg, "SALS")

```


```{r}
# Combine the plots into a 2x2 layout
combined_plot_2x2 <- ggarrange(
  go_dotplot, dotplot_alz, dotplot_c9als, dotplot_sals,
  labels = c("A", "B", "C", "D"),  # Add labels to indicate each plot
  ncol = 2, nrow = 2,  # Arrange in 2 columns and 2 rows
  common.legend = FALSE  # Each plot has its own legend
)

# Save the combined plot as a high-quality PNG for publication
ggsave("/Users/rutger/Desktop/Bureaublad/School/Master/Stage/Results/ThesisFigures/combined_GOBPandKEGG2x2_plot.png", 
       combined_plot_2x2, width = 14, height = 14, dpi = 600)
```




